---
title: "EDA"
author: "Nora Lee Reinhardt"
date: "`r Sys.Date()`"
output: html_document
---

This datafile is about the success of math students based on a set of input variables. The binary predictors are school, sex, address, famsize, Pstatus, schoolsup, famsup, paid, activities, nursery, higher, internet, romantic. The categorical predictors are Medu, Fedu, Mjob, Fjob, reason, guardian, traveltime, studytime. The numeric variables are failures, absences. The target variable is G4 which is a categorical variable to indicate a student’s success. There are 31 predictors and 273 values of each. There are no NAs.

```{r}
library(tidyverse)

mathscores <- read.csv("MathScores.csv")

view(head(mathscores))


```


```{r}
summary(mathscores)
```


```{r}
summary(mathscores$age)

mathscores %>% 
  ggplot(aes(x = age)) + 
  geom_bar(fill = "purple") +
  ggtitle('Age of Students')
```
The age of students ranges from 15 to 22 and the distribution is right skewed.
```{r}
summary(mathscores$failures)


mathscores %>% 
  ggplot(aes(x = failures)) + 
  geom_bar(fill = "hotpink") +
  ggtitle('Number of Failures')
```
The majority of students have 0 failures. Less than 50 students have one failure. 

```{r}
summary(mathscores$absences)

mathscores %>% 
  ggplot(aes(x = absences)) + 
  geom_bar(fill = "limegreen") +
  ggtitle('Number of Absences')

```
This distribution of absences is positively skewed. There are several outliers with high numbers of absences but the most common number of absences is 0.
```{r}
table(mathscores$sex)

mathscores %>% 
  ggplot(aes(x = sex)) + 
  geom_bar(fill = "skyblue") +
  ggtitle('Number of Sexes')

```
There are more female students in the data set than males. About 53.8% of the students are female.
```{r}
table(mathscores$address)


mathscores %>% 
  ggplot(aes(x = address)) + 
  geom_bar(fill = "lavender") +
  ggtitle('Number of adress types')
```
There are more students with urban addresses than rural addresses. Only 20.1% of the students have rural addresses.
```{r}
table(mathscores$school)

mathscores %>% 
  ggplot(aes(x = school)) + 
  geom_bar(fill = "blue") +
  ggtitle('Number of school types')

```
The majority of students attend school A. Only about 11.4% of the students are from school B.
```{r}
table(mathscores$reason)

mathscores %>% 
  ggplot(aes(x = reason)) + 
  geom_bar(fill = "yellow") +
  ggtitle('Reasons for choosing school')

```
The most popular reason for choosing a school was the courses offered and the reputation and it being close to home had similar results to eachother.
```{r}
table(mathscores$activities)

mathscores %>% 
  ggplot(aes(x = activities)) + 
  geom_bar(fill = "red") +
  ggtitle('extracurricular activities?')
```
The number of students in activities is close to the number of students not in activities. There is one more student who is not in activities.
```{r}
table(mathscores$schoolsup)

mathscores %>% 
  ggplot(aes(x = schoolsup)) + 
  geom_bar(fill = "brown") +
  ggtitle('Extra School Support?')
```
The majority of students do not have addition educational support. 
```{r}
table(mathscores$famsup)

mathscores %>% 
  ggplot(aes(x = famsup)) + 
  geom_bar(fill = "magenta") +
  ggtitle('Family School Support?')
```

```{r}
table(mathscores$paid)

mathscores %>% 
  ggplot(aes(x = paid)) + 
  geom_bar(fill = "cyan3") +
  ggtitle('Extra Paid Classes?')

```

```{r}
table(mathscores$traveltime)

mathscores %>% 
  ggplot(aes(x = traveltime)) + 
  geom_bar(fill = "orange") +
  ggtitle('Travel Time:')
```
```{r}
table(mathscores$studytime)


mathscores %>% 
  ggplot(aes(x = studytime)) +
  geom_bar(fill = "yellowgreen") + 
  ggtitle('Weekly Study Time')
```

```{r}
table(mathscores$freetime)

mathscores %>% 
  ggplot(aes(x = freetime)) + 
  geom_bar(fill = "lightsalmon") +
  ggtitle('Free Time after School')
```

